---
title: "Using Mapview"
---

The data is stored in `water_insecurity.rdata` in this folder. 

It contains the following data frames

- `fulldata` - both years (2022/2023) data as an `sf` `data.frame`
- `water_insecurity_2022` and `water_insecurity_2023` - data split by year
- `adi_2022` and `adi_2023` - Area deprivation index calculations split by year
- `raw_data_2022` and `raw_data_2022` - the data from ACS that was used to derive the ADI.

```{r}
library(mapview)
library(leaflet)
library(plainview)
library(leafsync)
library(tidyverse)
load("water_insecurity.rdata")
```

```{r}
head(fulldata)
```

We can use the `mapview` package to start exploring the dataset. We can view all rows of the data:

```{r}
mapview(fulldata)
```

We can filter by state:

```{r}
wa_counties <- fulldata |>
  dplyr::filter(state == "Washington") |>
  dplyr::filter(year == 2022)
```


We can visualize two variables as different layers by adding them together. Try clicking on the different layers (use the layer button in the top left).

```{r}
m1 <- mapview(wa_counties, zcol=c("adi")) 
m2 <- mapview(wa_counties, zcol=c("percent_lacking_plumbing"))

m1 + m2
```

Or we can sync two maps:

```{r}
sync(m1, m2)

```